\plaintitle

### 2016-12-13 drafts/labor/introduction.tex
cleaned refs
rebuttal init
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

\section{Introduction}\label{sec:introduction}

\topic{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
the constituent parts of larger tasks.}
This framing of work into de--contextualized, modular blocks
enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in the work whenever their schedules allow,
with little to no awareness of the broader context of the work, and 
with (often) fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:
\begin{inlinelist}
\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.
\end{inlinelist}
The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
like AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).

\topic{As on--demand work has grown far beyond information work, it has given rise to an increasingly complicated and conflicted culture amongst both the workers who enable it and the researchers who study it.}
\citeauthor{howe2008crowdsourcing} first described crowdsourcing in general terms as
``outsourcing [work] to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
However, for years its instantiation was limited to the utilization of
human intelligence to process data, participate in scientific studies, and perform information work~\cite{CrowdsourcingUserStudies,movieSummarizationWu,
      yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
More recently, crowdsourcing of physically embodied work
--- driving and cleaning, for instance ---
has become a focus for on--demand labor markets~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
This growth prompted efforts to understand not just the work, but also the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the identification of the sociality of gig work,
and the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
Other work has focused on the outcomes of this frustration,
reflecting on the resistance workers express against digitally mediated labor markets~\cite{uberAlgorithm,dynamo}.

@@

\subsection{Piecework as a lens to understand crowdsourcing}
% MSB Cutting because it seems like it's just attacking the literature unnecessarily...
% ----
%This large and growing body of research has conversed
%to varying degrees with labor scholarship,
% ...
%To the extent that the research community has explored the boundaries of crowd work,
%it nevertheless has not reflected on the underlying mechanisms determining those boundaries.
%Similarly, work directed toward the decomposition of tasks has begun to appreciate the limits of decomposition,
%but on the whole it hasn't considered the relationship between the decomposition of work and the complexity of work.
% \ali{Mixed feelings about this because I think I'll need to go back into decomposition and
% give a whole thing on \citeauthor{marx2012economic}
% if I bring this up as a real failing of the crowd work research to date.
% Thoughts?}
%While researchers are quickly picking up on the importance of
%the relationships between workers in these platforms,
%this research seems to be unfolding without the benefit of any of
%the inter--personal labor research that has been informing modern work for the better part of a century.
% \ali{Overall: too long? too vague? overstating the shortcomings?}
In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework}: %\namerefl{sec:pieceworkArc} ---
a work and payment structure which breaks tasks down into standalone contracts,
wherein payment is made for work output, rather than for time.
Piecework as a metaphor for crowd work is not new.
Indeed,
\citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
referenced crowd work as piecework briefly
as a loose analogy to the form of work emerging at the time~\cite{crowdworkFuture}.
But more than this,
the framing of on--demand labor as a re--instantiation of piecework
gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirically based research.

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by looking for patterns of behavior that the corresponding literature predicts
on this basis, we can
\begin{inlinelist}
  \item make sense of the phenomena so far as part of a much larger series of interrelated events;
  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today;
        and finally,
  \item to the extent that history repeats itself, offer predictions of what on--demand work researchers,
        and workers themselves,
        might expect to see on the horizon.
\end{inlinelist}
For example, we will draw on the piecework literature studying task decomposition, which was historically limited in scope by technological limits in measurement and instrumentation, and leverage that understanding to suggest how modern web technology reconfigures this measurement limit for on--demand crowd work.
%we draw upon the piecework literature to focus the research crowd work researchers have done on the complexity of work,
%and come up with theoretically grounded mechanisms explaining the limitations researchers have struggled to overcome thus far.
% \msb{what I think would really help this paragraph is a single great example, in one or two sentences, of how this kind of framing would help. ``For example, we draw on ... to explain ...''. Reader goes AHA I see why you're doing this now!}

We organize this paper as follows: we first review the definition and historical arc of piecework to lay  groundwork and make clear the analogy to on--demand work (which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature). Then, we interrogate the three major research questions above using the lens of piecework. 
% We will look at a broad range of cases under a number of major themes \msb{a broad range under a number of themes is an empty calorie sentence. get specific.}
% we propose as broadly describing the types of research being done in crowd work
% and more generally in what we argue is contemporary piecework.
%We will look at the above questions in crowd work through the lens
%first of the body of research made by crowd work researchers,
%and then of the body of work piecework researchers have offered.
%After validating this lens as a way of reasoning about crowd work and on--demand labor as a whole,
We will identify similarities and differences between piecework as historically understood and
on--demand work as we experience it today.
Finally, we will make predictions of future developments based on how those similarities and differences influenced piecework.
% we will use this perspective to suggest areas of research worth anticipating,
% and developments we should expect to see in the maturation of digitally mediated work.
Finally, we will offer implications for researchers and practitioners based on our results.

>>>>

\section{Introduction}\label{sec:introduction}
\topic{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
the constituent parts of larger tasks.}
This framing of work into de--contextualized, modular blocks
enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in the work whenever their schedules allow,
with little to no awareness of the broader context of the work, and 
with (often) fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:
\begin{inlinelist}
\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.~
\end{inlinelist}
The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
like AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).

\topic{As on--demand work has grown far beyond information work, it has given rise to an increasingly complex, conflicted culture amongst both the workers who enable it and the researchers who study it.}
\citeauthor{howe2008crowdsourcing} first described crowdsourcing broadly as
``outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
However, for years its instantiation was limited to the utilization of
human intelligence to process data, participate in scientific studies, and perform information work~\cite{CrowdsourcingUserStudies,movieSummarizationWu,
      yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
More recently, crowdsourcing of physically embodied work
--- driving and cleaning, for instance ---
has become a focus for on--demand labor markets~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
This growth prompted efforts to understand not just the work, but also the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the identification of the sociality of gig work,
and the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
Other work has focused on the outcomes of this frustration,
reflecting on the resistance workers express against digitally mediated labor markets~\cite{uberAlgorithm,dynamo}.

@@

\subsection{Piecework as a lens to understand crowdsourcing}
In this paper, we offer a framing for on--demand work as a contemporary instantiation of \textit{piecework}: %\namerefl{sec:pieceworkArc} ---
a work and payment structure which breaks tasks down into standalone contracts,
wherein payment is made for work output, rather than for time.
Piecework as a metaphor for crowd work is not new.
Indeed,
\citeauthor{crowdworkFuture} in \citeyear{crowdworkFuture}
referenced crowd work as piecework briefly
as a loose analogy to the form of work emerging at the time~\cite{crowdworkFuture}.
But more than this,
the framing of on--demand labor as a re--instantiation of piecework
gives us years of historical material to make sense of this new form of work, and allows us to reflect on--demand work through a mature theoretical lens, informed by decades of rigorous, empirically based research.

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by looking for patterns of behavior that the corresponding literature predicts
on this basis, we can
\begin{inlinelist}
  \item make sense of the phenomena so far as part of a much larger series of interrelated events;
  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today;
        and finally,
  \item to the extent that history repeats itself, offer predictions of what on--demand work researchers,
        and workers themselves,
        might expect to see on the horizon.
\end{inlinelist}
For example, we will draw on the piecework literature studying task decomposition, which was historically limited in scope by technological limits in measurement and instrumentation, and leverage that understanding to suggest how modern web technology reconfigures this measurement limit for on--demand crowd work.

We organize this paper as follows: we first review the definition and historical arc of piecework to lay  groundwork and make clear the analogy to on--demand work (which we will refer to as \textit{crowd work} subsequently, for consistency with prior literature). Then, we interrogate the three major research questions above using the lens of piecework. 
We will identify similarities and differences between piecework as historically understood and
on--demand work as we experience it today.
Finally, we will make predictions of future developments based on how those similarities and differences influenced piecework.
Finally, we will offer implications for researchers and practitioners based on our results.

### 2016-12-13 drafts/labor/piecework_lit.tex
piecework literature review section mostly updated. the various research questions might still be a bit of a mess

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it; and
  \item trace the rise and fall of piecework at a very high level,
        identifying key figures and ideas during this time;
  %        and finally
  % \item look at the fall of piecework, such as it was,
  %       considering in particular
  %       the factors that may have led to piecework's
  %       eventual demise in the American and European labor markets.
        % \msb{did it demise? there are sweatshops around the world today}.
        % \ali{is it okay to corner it off like this (by just being about the US/EU)?}
\end{inlinelist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

% While piecework has been studied from myriad perspectives through the lenses of
% political science,
% economics (both micro-- and macro--), and
% a plethora of other fields
% \ali{should I cite examples of all of these?}, for the sake of brevity
% we will discuss the history of piecework in the contexts of the three questions we brought up with crowd work:
% ``\namerefl{sec:complexity}'',
% ``\namerefl{sec:decomposition}'', and 
% ``\namerefl{sec:relationships}''.

@@

\topic{We see this sense of independence regardless of the time, locale, and industry.}
\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match girls''
--- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
Of particular interest was their independent nature, via their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia,
finding similar assertions of the freedom compositors of newspapers experienced as pieceworkers:
``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
From these accounts we identify a sense of independence and autonomy that resonates across decades, industries, and locales where piecework is found.
% We'll problematize this supposed advantage as we trace the history of piecework,
% but for now we can say that piecework affords
% independence and some sense of autonomy
% new to people in the working class.

\topic{Piecework opened the door for people who previously couldn't participate in the labor market --- for example due to lack of training --- to do so, and to acquire job skills incrementally.}
For example, women could receive training in narrow subsets of the general body of skills, enabling them to act in capacities similar to what conventional (male) apprentices would undertake~\cite{hart2013rise}. 
% offer another series of compelling insights toward
% the question of the features that sprout from piecework.}
% In their reflection on the features endemic to piecework in the 1930s,
% which they describe as the ``heyday'' of piecework's prominence;
% among them were the following:
% \begin{inlinelist}
%   \item ``female workers who generally had less training'' had to be trained in
%         narrower subsets of the general body of skills that
%         conventional (male) apprentices would undertake, and
In addition, workers with specific slices of skills could be matched to suitable tasks.
Workers without conventional training
--- like women, who had no such opportunities
to engage in engineering and metalworking apprenticeships as men did ---
could be trained very narrowly on a very tightly constrained task,
demonstrate proficiency, and become experts in their own ways.

In summary, piecework:
\begin{enumerate}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models;
  \item afforded workers freedom in when and how much to work; and
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{enumerate}
% \ali{woah i thought this was a comment; is it okay as is?}

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition. 
First, platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well. 
% Second, workers on these platforms have significant freedom in choosing which tasks to work on, and when to log on and perform work~\cite{martin2014being,whyWouldAnyoneBrewer}. 
% Finally, computational workflows have successfully scaffolded many complex tasks into a series of small microtasks that can be performed without expertise.
%  \subsection{Piecework and crowd work}
% \msb{I assume this is forthcoming?}
% Using the definition of piecework that we came up with earlier, we argue that
% crowd work is fundamentally an instantiation of piecework, and
% that we can more precisely anticipate the answers to the open research questions we discussed earlier.
% We'll show that the dimensions of crowd work that the broader HCI community has been studying
% align with the history of piecework, and that this can greatly inform predictions about the future of crowd work.
% \subsection{From piecework to on--demand work}
% Crowd work and gig work are fundamentally an instantion of piecework.
% First, workers on platforms such as Mechanical Turk and Uber are generally incentivized by unit of work, even if some may be offered an hourly base salary as well.
Second, workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
Third, system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.

% \ali{REMOVED A BUNCH OF STUFF FROM HERE ONWARD}

\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % yay rhyming?
\topic{In this section we will offer a brief overview of the history of piecework;
this should not be mistaken for a comprehensive background.}
Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, it will frame piecework in the contexts of
the early days of the Industrial Revolution,
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework.

@@

% While this approach didn't become the same economic powerhouse as later examples would,
% Airy and others arguably found the kernel of insight that we pursue throughout this discussion:
% determining the extent to which work can be decomposed, and
% finding the limits of complexity of that decomposed work.
% Airy found that he could train youths in elementary mathematics
% to complete the majority of the calculations he would otherwise have had to solve on his own,
% and that the greater body of work could ultimately be completed sooner
% if he arranged his work appropriately.
% \msb{After reading this paragraph, I don't know what it's supposed to be teaching me.
% What I got out of it is that a bunch of people did piecework, but
% I don't know why or in fact why these are different than the sources we cite earlier.
% Can you hone it?}
% \ali{I wanted that paragraph to be about the rising popularity and application of piecework
% (especially as it approached its ``heyday'' \cite{hart2013rise}),
% coming from humble beginnings as it found its footing.
% Given that intent\dots\\
% Should I \textbf{refactor} or \textbf{rewrite}?}

% \ali{\sout{I'm thinking of
% dropping most of the last paragraph and merging it with the next one. Thoughts?}
% I moved it to \nameref{sec:complexity}}

% \ali{Where do I want to take this? Where are we now?
% \begin{enumerate}
%   \item Piecework started in the \nth{19} century (above). [DONE]
%   \item Piecework went on to be a pretty cool thing in textiles and farming at the turn of the \nth{20} century. [DONE?]
%   \item Piecework went gangbusters through the first half of the century, especially through the Second World War. []
%   \item Piecework died pretty quickly (below). [DONE]
% \end{enumerate}
% Sound good?}

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took off on in farm work \cite{hughRaynbirdTaskWork},
in textiles \cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads \cite{Brown01041990}, and 
elsewhere in manufacturing \cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\citeauthor{hart2013rise} point out that the Second World War,
which called millions of Americans to military service,
necessitated the rapid training and employment of
a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.
% With little time and virtually none of the training that was conventionally expected of metalworkers and other industrial work
% (e.g. years--long apprenticeships),
% tasks had to be redesigned and workers trained for these dramatically unorthodox task designs
\ali{I'd like to bring the Rosie the Riveter anecdote in here but it's not feeling natural\dots}
% \cite{honey1985creating,davies2014origins}.

The early proliferation of manual piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
Despite this, workers' means were mostly ignored,
leading to frustration over poor working and living conditions
(famously documented by
% \ali{??? Why is this here? Can it be moved?}
% {As increasing attention revealed problems in piecework as it related to workers,
% workers themselves began to speak out about their frustration with this new regime.}
\citeauthor{riisOtherSideLives})~\cite{riisOtherSideLives}.
This led to industry organizations representing
railway workers, mechanical engineers, and others beginning to speak out on pieceworkers' behalf~\cite{american1921problem,richards1904anything}.
% Nevertheless, piecework continued to permeate low--skilled labor.
% \ali{How does this tie in with the rest of the narrative? Can I move this around? Does it break everything to take it out of chronological order?}

@@

\onlyinsubfile{
  \printbibliography
  % \clearpage
  }

>>>>

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{inlinelist}
  \item define ``piecework'' as researchers in the topic understood it; and
  \item trace the rise and fall of piecework at a very high level,
        identifying key figures and ideas during this time;
\end{inlinelist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

@@

\topic{We see this sense of independence regardless of the time, locale, and industry.}
\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
--- young women paid by piecework to assemble matchsticks generally in the late \nth{19} century.
Of particular interest was their independent nature, via their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} documents piecework from 1850--1930 in Australia,
finding similar assertions of the freedom compositors of newspapers experienced as pieceworkers:
``If a piece--work compositor who held a `frame' decided that he did not want to work on a particular day or night,
the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
From these accounts we identify a sense of independence and autonomy that resonates across decades, industries, and locales where piecework is found.

\topic{Piecework opened the door for people who previously couldn't participate in the labor market --- for example due to lack of training --- to do so, and to acquire job skills incrementally.}
For example, women could receive training in narrow subsets of the general body of skills, enabling them to act in capacities similar to what conventional (male) apprentices would undertake~\cite{hart2013rise}.
In addition, workers with specific slices of skills could be matched to suitable tasks.
Workers without conventional training
--- like women, who had no such opportunities
to engage in engineering and metalworking apprenticeships as men did ---
could be trained very narrowly on a very tightly constrained task,
demonstrate proficiency, and become experts in their own ways.

In summary, piecework:
\begin{inlinelist}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models;
  \item afforded workers freedom in when and how much to work; and
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
\begin{Numberlist}
\item platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
\item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
\end{Numberlist}
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.

\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\topic{In this section we will offer a brief overview of the history of piecework;
this should not be mistaken for a comprehensive background.}
Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, it will frame piecework in the contexts of
the early days of the Industrial Revolution,
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework.

@@

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took off on in farm work~\cite{hughRaynbirdTaskWork},
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\citeauthor{hart2013rise} point out that the Second World War,
which called millions of Americans to military service,
necessitated the rapid training and employment of
a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.

The early growth of piecework led to discussion surrounding how best to manage pieceworkers~\cite{norton1900textile,clark1908cotton}.
Despite this, workers' needs were mostly ignored,
leading to frustration over poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives})~\cite{riisOtherSideLives}.
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

@@

\subsection{Case studies in piecework}
We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
\begin{numberlist}
  \item Airy's employment of \textit{human computers},
  \item domestic and farm work (in particular, the ``match--girls''), and
  \item railroad workers across the United States.
\end{numberlist}

\ali{I'm coming out of this with a sort of structure that \textit{loosely} describes piecework chronologically.
Strictly speaking, I'm talking about each of these topics
(let's call them
\textbf{human computation},
\textbf{domestic and farm work}, and
\textbf{industrial work}),
and some of these things overlap with each other to some extent, but
for the most part
Airy and the human computers came in the mid--\nth{19} century,
farm work and the ``domestic'' work (like making matches) came around at the turn of the \nth{20} century, and
the industrial work (railroad workers, the WWII war effort, and the rise of the labor union movement) unfolded from the early \nth{20} century onward.}

\subsubsection{Airy's Computers}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
% piecework, found in George Airy's creation of
% the British Nautical Almanac.}
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.

\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
For one thing,
Airy had set a policy of firing his computers once they reached age 23.
This practice ensured two outcomes that arguably disfavored workers.
\begin{Numberlist}
\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
\item it limited workers' ability to organize
      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
\end{Numberlist}
% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}

\subsubsection{Domestic and Farmhand Labor}
\topic{When piecework entered the American economy, it was not used for complex work.}
One reason for low complexity was workers' skills:
it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted~\cite{hart2013rise}.
So, initially piecework in the United States arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
The complexity of the work remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
However, writers of the time focused their attention on
wage~\cite{burton1899commercial} and
management regimes~\cite{norton1900textile}
rather than training.
Mass manufacturing,
such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
flourished under piecework systems in densely populated cities.

\topic{Workers' relationships with employers quickly soured.}
The match--girls strike of 1888 was one of the earliest and most famous successful worker strikes,
and perhaps the beginning of ``militant trade unionism''~\cite{10.2307/3827491}.
As \citeauthor{weyer1894history} described,
``the match--girls' victory turned a new leaf in Trade Union annals''~\cite{weyer1894history}: in the 30 years after the match--girls strike,
the Trade Union Movement enrollment grew from 20\% of eligible workers to over 60\%.

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
In 1912, national coal miners effected an individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Women in the garment industry in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
% and it was this framing on work that
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor himself, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

\subsubsection{Railroad and other Industrial Workers}
\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
Scientific management argued that it was possible to measure work at
unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
As \citeauthor{Brown01041990} points out,
piecework most greatly benefits the instrumented measurement of workers,
% but certainly in Ford and Taylor's time,
but in \citeauthor{taylor1914principles}'s time highly instrumented,
automatic measurement of workers was all but impossible~\cite{Brown01041990}.
Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
as rudimentary guides on per--task compensation.
This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
railroad maintenance, automobile manufacturing, and later weapons manufacturing.
\ali{???? better word for ``weapons manufacturing''?}
% Instead, as a result,
% the distillation of work into smaller units ultimately
% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.

% \topic{With the formalization of scientific management,
% piecework became an important component of the mass manufacturing
% in American factories.}
% \topic{}

\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

\onlyinsubfile{
  \printbibliography
  }
